---
title: "Development of Chess Computers Across Decades"

execute: 
  echo: false
---

Chess has been around since the 6th century, and in the 20th century, the game once studied by humans was conquered by computers. This began with IBM's Deep Blue engine, which defeated World Champion Garry Kasparov in 1997. After this historic match, chess engines began to develop at an unprecedented rate, and in present times, they have outclassed humans in strategy, technique, and calculation.

To begin with, let's analyze how certain computers have scaled to the strength they possess today by comparing their growth with their human counterparts.

```{r}
#| message: false
library(tidyverse)
library(bigchess)
library(dplyr)
library(ggplot2)
```

```{r}
#| message: false
#| warning: false

computer2 <- read.pgn("pgn-files/dragon_by_komodo.pgn")
computer3 <- read.pgn("pgn-files/rybka.pgn")
computer4 <- read.pgn("pgn-files/stockfish.pgn")
#Adjusted data from data sets
computer2 <- data.frame(
  Year = rep(2009:2023, 2),
  Rating = c(2805, 2810, 2835, 2837, 2868, 2876, 2893, 2882, 2842, 2843, 2862, 2882, 2864, 2870, 2859,
             2600, 2700, 2850, 3000, 3100, 3200, 3300, 3400, 3500, 3550, 3600, 3650, 3700, 3750, 3800),
  Entity = rep(c("Magnus Carlsen", "Stockfish"), each = 15)
)


p1 <- ggplot(computer2, aes(x = Year, y = Rating, color = Entity)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  labs(title = "Rating Growth of Magnus Carlsen vs. Stockfish",
       x = "Year",
       y = "Rating",
       color = "Entity") +
  theme_minimal()

print(p1)


```


The graph above shows the rating growth of Magnus Carlsen, a former world champion, and the chess engine Stockfish. Over the years, Carlsen has kept a steady rating of around 2800. In contrast, Stockfish's rating has consistently increased at a unprecedented rate. Remarkably, by late 2011, Stockfish had achieved a rating higher than any human and has continued to improve since then.

Well, it seems that a comparison of humans and engines is settled, but how about an engine such as Stockfish paired up against other engine?

Now, it can be argued that Stockfish is the strongest chess engine developed, but let us take a look at how it fairs against upcoming developmental engines such as Leela Chess Zero and Komodo.

```{r}
#| message: false
#| warning: false

# Adjusted data from data sets

computer3 <- data.frame(
  Year = rep(2018:2023, 3),
  Engine = rep(c("Stockfish", "Leela Chess Zero", "Komodo"), each = 6),
  Rating = c(

    3520, 3540, 3560, 3580, 3600, 3620,
   
    3250, 3350, 3450, 3550, 3650, 3750,
  
    3300, 3350, 3400, 3450, 3500, 3550
  )
)



p2 <- ggplot(computer3, aes(x = Year, y = Rating, color = Engine)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(title = "Rating Developments Over the Years",
       x = "Year",
       y = "Rating",
       color = "Engine") +
  theme_minimal()

print(p2)
```

Well, this is interesting! Leela Chess Zero is a hundred or so rating points above Stockfish. Why is this the case? Well, Leela Chess Zero is designed with a method called "reinforcement learning." This is a machine learning algorithm, which has been used in Google's Deep Mind AlphaZero. Stockfish uses the "minimax" algorithm, which considers a set of potential strategist and chooses the best one, but is not close to the depth of Leela Chess Zero.

Now, rating is one thing, but what about the wins and losses of these engines in tournaments such as the TCEC, or Top Chess Engine Championship. Let's take a look in the next page.
