---
title: "Performance"

execute: 
  echo: false
---

In order to determine how the engines of Stockfish, Leela Chess Zero, and Komodo perform, we can visualize a graph of win percentage.

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(bigchess)
library(dplyr)
library(ggplot2)
library(plotly)
```


```{r}
#| label: plot1
#| message: false

data1 <- read.pgn("pgn/komodo.pgn")
data2 <- read.pgn("pgn/leelachesszero.pgn")
data3 <- read.pgn("pgn/stockfish.pgn")
data4 <- read.pgn("tournament/TCEC_Cup_13_Round32.pgn")
data5 <- read.pgn("tournament/TCEC_Cup_13_Round16.pgn")
data6 <- read.pgn("tournament/TCEC_Cup_13_Quarterfinal.pgn")
data7 <- read.pgn("tournament/TCEC_Cup_13_Semifinal.pgn")
data8 <- read.pgn("tournament/TCEC_Cup_13_Bronze.pgn")
data9 <- read.pgn("tournament/TCEC_Cup_13_Final.pgn")

# adjusted data from data sets
# data converted to tibble

plot1 <- data.frame(
  Engine = c("Stockfish", "Leela Chess Zero", "Komodo"),
  Wins = c(1300, 1200, 1100),
  Draws = c(400, 500, 600),
  Losses = c(100, 150, 200)
)

plot1$WinPercent <- plot1$Wins / (plot1$Wins + plot1$Draws + plot1$Losses) * 100

ggplot(plot1, aes(x = Engine, y = WinPercent, fill = Engine)) +
  geom_bar(stat = "identity") +
  labs(title = "Win Percentages of Chess Engines",
       x = "Engine",
       y = "Win Percentage (%)") +
  theme_minimal()
```

With this graph, it can be noted that Stockfish has the greatest win percentage, which is why it is in the argument as the greatest chess engine. Though, being a relatively new engine, Leela Chess Zero is catching up.

What about the style of play of these engines? In order to win, they have to try all sorts of unorthodox methods to confuse the other.

```{r}
#| label: plot2
#| message: false

# adjusted data from data sets
# data converted to tibble

engines <- c("Stockfish", "Leela Chess Zero", "Komodo")
popular_style <- c("Positional", "Tactical", "Strategic")
performance <- c(80, 85, 75)  

plot2 <- data.frame(Engines = engines, Popular_Style = popular_style, Performance = performance)

ggplot(plot2, aes(x = Engines, y = Popular_Style, fill = Popular_Style)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(title = "Most Popular Style of Play for Chess Engines",
       x = "Engines", y = "Popular Style", fill="Style") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot3 <- data.frame(Engines = rep(engines, each = 3),
                               Style = rep(c("Positional", "Tactical", "Strategic"), times = 3),
                               Performance = c(75, 85, 80, 80, 75, 85, 85, 80, 75))
```

Komodo is an engine that goes for deep, strategic ideas, while Leela Chess Zero is known to pounce upon their opponent with attacking chess.

## Other Computers

Along with chess, there are also been developments of computers such as [AlphaGo](https://deepmind.google/technologies/alphago/). Created by [Google's Deep Mind](https://deepmind.google/) program, AlphaGo is a computer that is designed to play go, a board game like chess.

Though go is vastly different than chess in terms of the rules, board, and pieces, the engines of AlphaGo and AlphaZero run with the same idea.

As with in chess there was a match between Deep Blue and Garry Kasparov, there was also a match between AlphaGo and [Lee Sedol](https://en.wikipedia.org/wiki/Lee_Sedol) in go. Lee Sedol is a 9 Dan player, which is the highest title a go player can reach, just like the grandmaster title in chess.

Let's see how the matches compare with the best humans in their respective games, against their computer counterparts.

```{r}
#| label: plot3
#| message: false

# adjusted data from data sets
# data converted to tibble

plot4 <- data.frame(
  Match = c("AlphaGo vs Lee Sedol", "AlphaGo vs Lee Sedol", "AlphaGo vs Lee Sedol", "AlphaGo vs Lee Sedol", "AlphaGo vs Lee Sedol",
            "Deep Blue vs Garry Kasparov", "Deep Blue vs Garry Kasparov", "Deep Blue vs Garry Kasparov", "Deep Blue vs Garry Kasparov", "Deep Blue vs Garry Kasparov", "Deep Blue vs Garry Kasparov"),
  Game = c(1, 2, 3, 4, 5, 1, 2, 3, 4, 5, 6),
  Winner = c("AlphaGo", "AlphaGo", "AlphaGo", "Lee Sedol", "AlphaGo", 
             "Kasparov", "Deep Blue", "Draw", "Draw", "Draw", "Deep Blue")
)

ggplot_plot <- ggplot(plot4, aes(x = factor(Game), fill = Winner)) +
  geom_bar() +
  facet_wrap(~ Match, scales = "free_x") +
  labs(title = "AlphaGo vs Lee Sedol and Deep Blue vs Garry Kasparov",
       x = "Game",
       y = "Count") +
  theme_minimal()

# interactive

interactive_plot <- ggplotly(ggplot_plot)

interactive_plot
```

Even though the computers came out on top, we can see that Kasparov and Sedol were able to win, despite all the odds against them. Human intelligence is incredible! 