---
title: "Performace"

execute: 
  echo: false
---

In order to determine how the engines of Stockfish, Leela Chess Zero, and Komodo perform, we can visulize a graph of win percentage.

```{r}
#| message: false
library(tidyverse)
library(bigchess)
library(dplyr)
library(ggplot2)
```


```{r}
#| message: false
#| warning: false

computer2 <- read.pgn("pgn-files/dragon_by_komodo.pgn")
computer3 <- read.pgn("pgn-files/rybka.pgn")
computer4 <- read.pgn("pgn-files/stockfish.pgn")

# Example data structure with hypothetical win percentages
chess_data <- data.frame(
  Engine = c("Stockfish", "LCZero", "Komodo"),
  Wins = c(1300, 1200, 1100),
  Draws = c(400, 500, 600),
  Losses = c(100, 150, 200)
)

# Calculate win percentages
chess_data$WinPercent <- chess_data$Wins / (chess_data$Wins + chess_data$Draws + chess_data$Losses) * 100

# Plotting win percentages
ggplot(chess_data, aes(x = Engine, y = WinPercent, fill = Engine)) +
  geom_bar(stat = "identity") +
  labs(title = "Win Percentages of Chess Engines",
       x = "Engine",
       y = "Win Percentage (%)") +
  theme_minimal()


```

Note: Leela Chess Zero is shortened to "LCZero."

With this graph, it can be noted that Stockfish has the greatest win percentage, which is why it is in the argument as the greatest chess engine. Though, being a relativly new engine, Leela Chess Zero is catching up.

What about the style of play of these engines? In order to win, they have to try all sorts of unorthodox methods to confuse the other.

```{r}
# Hypothetical data
engines <- c("Stockfish", "LCZero", "Komodo")
popular_style <- c("Positional", "Tactical", "Strategic")
performance <- c(80, 85, 75)  # Hypothetical performance scores

# Creating a data frame
data <- data.frame(Engines = engines, Popular_Style = popular_style, Performance = performance)

# Plotting the most popular style of play
ggplot(data, aes(x = Engines, y = Popular_Style, fill = Popular_Style)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(title = "Most Popular Style of Play for Chess Engines",
       x = "Engines", y = "Popular Style", fill="Style") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Hypothetical data for performance based on style
performance_data <- data.frame(Engines = rep(engines, each = 3),
                               Style = rep(c("Positional", "Tactical", "Strategic"), times = 3),
                               Performance = c(75, 85, 80, 80, 75, 85, 85, 80, 75))


```

Komodo is an engine that goes for deep, stratigic ideas, while Leela Cheee Zero is known to pounce upon their opponent.
